---
title: "Five Parameters"
author: "Sophia Yao"
format: html
execute:
  echo: false
---

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(primer.data)
library(brms)
library(tidybayes)
library(gtsummary)
```

```{r}
#| label: model
#| cache: true

df1 <- governors |>
  select(last_name, year, state, sex, lived_after, election_age)

fit_all <- brm(data = df1, formula = lived_after ~ sex*election_age, silent = 2, refresh = 0, seed = 13)
```

```{r}
ndata <- tibble(sex = c("Male", "Female"), 
       election_age = 50)

fit_all |>
  add_epred_draws(newdata = ndata) |>
  ggplot(aes(.epred, fill = sex)) +
  geom_histogram(aes(y = after_stat(count/sum(count)),
                alpha = 0.5, 
                bins = 100, 
                position = "identity")) + 
  scale_x_continuous(labels = scales::number_format(accuracy = 1)) + 
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) + 
  labs(title = "Posterior for Expected Years Lived Post-Election",
      subtitle = "Male candidates live longer",
      x = "Years",
      y = "Probability",
      colour = "sex")
```

Using data from all deceased gubernatorial candidates in the United States between 1945 and 2012, we seek to forecast candidate longevity post-election. It may not be representative because only the two candidates with the most votes are included in the data set. We built a model using brms in R to explore the relationship between the candidate's sex, their age at election, and how these factors interact. Our model indicates that male candidates generally live longer post-election than female candidates. At an election age of 50, males also live longer than female candidates.

$$ 
lived\_after_i =  \beta_0 + \beta_1 male_i + \beta_2 c\_election\_age_i + 
\\ \beta_3 male_i *  c\_election\_age_i + \epsilon_i 
$$
